Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-09-08 GWLDCII.tex
edits and comments to sections 3 and 4, paper 2

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$ under a counterclockwise orderin. That is, $p_1$ is closest to the vertex $e$, $p_i$ is positioned counterclockwise to $p_{i+1}$ and $p_r$ is the clockwise most propagator on this edge (closest to vertex $e+1$). Define the polynomial
\[ R_e =  x_{p_1,e+1} \prod_{j= 1}^{r-1} \left((x_{p_j,e} x_{p_{j+1},e+1} - x_{p_{j+1},e} x_{p_{j},e+1} ) \right) x_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = x_{p,e} x_{p,e+1}$. If $r=0$, set $R_e = 1$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}

@@

The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}. An worked example of Algorithm~\ref{alg:put GN on WLD} is given in Example~\ref{eg:apply GN alg}.

@@

\begin{enumerate}
\item Fix a vertex $i \in [n]$, the {\bf starting vertex}. Set $j:=i$ and $I_i^W = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps:
\begin{enumerate}
\item {\em Step $j$ for starting vertex $i$}: If $\Prop(j) = \emptyset$, do nothing. Else, let $p \in \Prop(j)$ be the clockwise most propagator supported on $j$. Write $I_i^W = I_i^W\cup \{j\}$, and delete propagator $p$ from the diagram.
\item Increment $j$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

Informally: one starts at $i$ and moves counterclockwise around the vertices of $W$, at each step removing the clockwise-most propagator supported on that vertex (if it exists). The set $I_i$ lists the vertices at which a propagator was deleted.

If the algorithm assigns vertex $j$ to propagator $p$ from starting vertex $i$, we say that $p$ \emph{contributes} $j$ to $I_i$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. For each $i \in [n]$, define a function $I_i^W : \cP \longrightarrow [n]$ by
\[I_i^W(p) := \text{the vertex label that $p$ contributes to $I_i$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$. 
\end{dfn}

For both $I_i^W$ the set and the function, when the diagram $W$ is clear from context, we drop the superscript and simply write $I_i$.

% Sian deleted the propagator ordering for now: since we only use it in one proof, we can make it a local definition.
% Obviously if we find ourselves using it elsewhere we can bring this definition back. (If so, uncomment the line in the following example as well.)

%Each $I_i$ also imposes an ordering on the propagators, determined by when the propagator contributes to the sequence:
\begin{dfn} \label{proporderingdfn}
  There is an ordering on the propagators on $W = (\cP, [n])$ defined by $I_i$. This is the ordering in which they contribute to the algorithm as it constructs $I_i$.
%  $\cP = \{p_1, \ldots p_k\}$, where 
% \[I_i = \{I_i(p_1) <_i I_i(p_2) <_i \ldots <_i I_i(p_k)\}.\]
% (The superscript $W$ is omitted here to simplify the notation.)\note{Sian tweaked this slightly to emphasise the $<_i$ order, because standard convention in most papers is to write all GN terms in the 1-order.}
\end{dfn}

Since each propagator is removed by the algorithm after it contributes, each propagator contributed exactly once to $I_i$, making the above ordering well defined.
Note that different $I_*$ impose different orderings on the propagators. Furthermore, Corrollary \ref{GN alg well defined} shows that this ordering respects the ordering of vertices in $I_i$. That is, $I_i(p_j) <_i I_i(p_l)$ for all $j <l$.

@@

To construct $I_1$ we start at vertex 1, so we set $i=1$, $j = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $j = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, $r$ contributes at this step: $I_1(r)=1$. We let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $j = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators, $q$ contributes at this step: $I_1(q)=1$. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $j = 3$: $\Prop(3) = \{p\}$, so now $I_i(p) = 3$ and $I_1 = \{1,2,3\}$. The propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $j = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $j = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$ with $I_1(s) = 5$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. 
%So the propagators are ordered $r,q,p,s$ with respect to $I_1$, and 
$I_1$ viewed as a function is given by
\[I_1: \cP \longrightarrow [n] : \quad I_1(r) = 1,\quad I_1(q) = 2, \quad I_1(p) = 3, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid $M(W)$ associated to the Wilson loop diagram $W$ above, as we will prove in general shortly.
\end{eg}

%Notice that for each element in the Grassmann necklace, the propagators contribute in a particular order. It is sometimes useful to think of the propagators in this order. 
%\begin{dfn} \label{prop order dfn}
%Given a set $\cP$ of $k$ propagators defining a Wilson loop diagram $(\cP, [n])$, the Grassmann necklace element $I_i$ imposes an ordering on it $\cP^i = \{p_1 \ldots p_k\}$ where $p_m$ is the $m^{th}$ propagator to contribute to $I_i$. We write $\cP_m^i = \{p_1 \ldots p_m\}$ to be the first $m$ propagators that contribute to $I_i$. We drop the superscript if the Grassman necklace element is clear.
%\end{dfn}

@@

\begin{figure}
\[
\begin{tikzpicture}[baseline=(current bounding box.east)]
	\begin{scope}
	W' = \drawWLDfragment[12]{3}{0.4} 
	%\drawnumberspartial % use this to debug
	\newnode[left]{1}{i}
	\newnode[below]{4}{j}
	\newnode{5}{}
	\newnode{6}{}
	\newnode{7}{}
	\newnode{8}{}
	\newnode{9}{}
	\newnode{11}{}
	\newnode[right]{12}{l+1}
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{6.3}{8.6}{80}
		\node at (\zero + 7.5*\step:\radius*0.8) {\footnotesize $r$};
		\newprop[midway,left]{5.5}{-4}{}
		\node at (\zero + 3*\step:\radius*0.5) {\footnotesize $p$};
		\newprop[midway]{11.5}{-4}{}
		\node at (\zero + 12.2*\step:\radius*0.7) {\footnotesize $q$};
	\end{scope}
\end{scope}
\end{tikzpicture}
\qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.east)]
	V' = \begin{scope}
	\drawWLDfragment[12]{3}{0.4} 
	%\drawnumberspartial % use this to debug
	\newnode[left]{2}{m}
	\newnode[below]{4}{j}
	\newnode{5}{}
	\newnode{6}{}
	\newnode{7}{}
	\newnode{8}{}
	\newnode{9}{}
	\newnode{11}{}
	\newnode[right]{12}{l+1}
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{6.3}{8.6}{80}
		\node at (\zero + 7.5*\step:\radius*0.8) {\footnotesize $r$};
		\newprop[midway,left]{5.5}{19}{}
		\node at (\zero + 5.5*\step:\radius*0.5) {\footnotesize $s$};
		\newprop[midway]{11.5}{21}{}
		\node at (\zero + 11.8*\step:\radius*0.8) {\footnotesize $t$};
	\end{scope}
\end{scope}
\end{tikzpicture}
\]
\caption{$W'$ and $V'$ are locally the same between $j$ and $l+1$, so we will have $I_i^W\cap [j,l] = I_m^V\cap[j,l]$ even though $p \neq s$ and $q \neq t$.}\label{fig:locally identical}
\end{figure}

@@

With that in mind, we are ready to begin proving properties of the algorithm. We first show that the algorithm for $I_i$ can never reach the final vertex (in the $<_i$ order) in a propagator's support without having already assigned that propagator; in particular, this guarantees that the construction of $I_i$ always terminates in fewer than $n$ steps, as shown in Corrollary \ref{GN alg well defined}.

@@

	\end{scope}
		\end{scope}
	\end{tikzpicture}
\] 
\caption{The two types of configuration in which a propagator could (in theory) contribute its $<_i$-maximal support vertex to $I_i$.}
\label{fig:no fourth vertex}
\end{figure}
\note{Question: is the diagram clearer in this configuration (starting vertex $i$ matches up in both) or in the old configuration (propagator $p$ matches up)?} \sanote{I think it is fine as is}

In Case 1, \st{observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$.}{\color{violet} since $b+1<_i I_i(p)$ there must be some propagator, $q$ before $p$ in $\cP^i$ such that $I_i(q) = b+1$. By remark \ref{clockwise ordering rem}, $p$ is inside $q$ in the $<_b$ order. Specifically, $q$ must also end on the $b^{th}$ edge. By non-crossing and desnity arguements of admissibility $i-1$ cannot be in $V(q)$.} Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of Case 2. Therefore it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

So let $p = (a,b)$ with $a<_ib$ and $I_i(p) = b+1$, and choose $p$ such that $\big|[a+1,b]\big|$ is minimal amongst propagators of $W$ which contribute their last vertices and are in Case 2.

@@

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators and $n$ vertices. Then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ distinct vertices to each $I_i$ in $n$ steps.
\end{cor} 
\begin{proof}
Fix $i \in [n]$. It follows from the proof of Lemma~\ref{lem no fourth vertex} that in applying Algorithm~\ref{alg:put GN on WLD} to construct the set $I_i$, the algorithm can never reach the fourth vertex of a propagator's support (ordered with respect to $<_i$) without having already assigned that propagator. Therefore if the algorithm starts at vertex $i$, all propagators must have been eliminated by the time it gets to vertex $i-1$, ensuring that the algorithm is completed in less than one full circumnavigation of the diagram and that $I_i$ contains exactly $k$ distinct elements.
\end{proof}

@@

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. \hlfix{Note that $p$ has no effect on other propagators for starting vertices in this range,}{By rmk \ref{rmk algorithm locally same}? If so, the entire proof depends on this fact. We should state it prominently and move on.} so if $r \neq p$, then $I_w^W(r) = I_w^V(r)$. Meanwhile, the value of $I^{W}_w(p)$ depends on which step in the algorithm the propagator $q$ contributes a vertex in the diagram $V$, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then 
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority 
%$w \in [l, l+a-1]$} 
%I left this as a comment so we can add it in if necessary, but I don't think we need this level of precision in this sentence: we're just signposting what we're going to do in detail shortly.
of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$ (recall Remark~\ref{rmk algorithm locally same}).  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in Case 2 of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is a Grassmann necklace.
\end{thm}
\begin{proof}
%\note{Sian replaced ``counterclockwise'' with ``clockwise'' in a couple of places in this proof: someone please doublecheck?  Karen checked and it seems correct with Sian's changes.}
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. We prove that $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is equivalent to the definition of Grassmann necklace given in Definition~\ref{def:grassmann necklace}.

Suppose for a contradiction that there exists an admissible diagram for which there exists an $i$ with $m\in I_i\setminus \{i\}$ and $m \not\in I_{i+1}$.  Fix $n$.  Let $W$ be such a counterexample with the minimal number of propagators.

If $i \not\in I_i$, then there are no propagators supported on $i$ at all.  In this case it is clear that applying Algorithm~\ref{alg:put GN on WLD} at vertex $i$ and vertex $i+1$ produces exactly the same result, i.e. $I_{i+1} = I_i$, and so $(W, i, m)$ is not a counterexample at all.

@@

\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$.  If $p$ is not supported on $i+1$, then in building $I_{i+1}$ we will assign the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, that is \bas I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]\eas by Remark \ref{rmk algorithm locally same}.  If $p$ is supported on $i+1$, then $b= i+1$ and the interval $[i+1,b-1]$ is empty, so the above equality remains true, albeit vacuously.  Furthermore, by Lemma~\ref{vertex cyclic int lem} it must happen that $p$ contributes vertex $b$ to $I_{i+1}$, as otherwise $b$ would never be contributed by $p$.  Consequently, by Remark \ref{clockwise ordering rem}, in building $I_{i+1}$,  when the algorithm reaches vertex $b$ there cannot be any remaining unassigned propagators that are supported on $b$ and clockwise of $p$.  This is also true in when the algorithm constructing $I_i$ reaches $b$, \st{since the same propagators have been assigned beforehand.}\sanote{don't like same here}{\color{violet}since any propagators lying counterclockwise of $p$ on the edge $i-1$ must have their support in $[i-1, b]$. By Lemma \ref{no fourth vertex}, these propagators contribute before the algorithm reaches $b$. Furthermore, Remark  \ref{rmk algorithm locally same} implies that these propagators contribute the same vertices to both $I_i$ and $I_{i+1}$.} This also implies that $m\geq_i b+1$.

Let $W' = (\cP_{out}(p,<_i), [n])$, i.e. the diagram obtained from $W$ by removing both $p$ and all propagators inside of $p$.

>>>>

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$ under a counterclockwise ordering. That is, $p_1$ is closest to the vertex $e$, $p_i$ is positioned counterclockwise to $p_{i+1}$ and $p_r$ is the clockwise most propagator on this edge (closest to vertex $e+1$). Define the polynomial
\[ R_e =  x_{p_1,e+1} \prod_{j= 1}^{r-1} \left((x_{p_j,e} x_{p_{j+1},e+1} - x_{p_{j+1},e} x_{p_{j},e+1} ) \right) x_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = x_{p,e} x_{p,e+1}$. If $r=0$, set $R_e = 1$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}

@@

The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}. A worked example of Algorithm~\ref{alg:put GN on WLD} is given in Example~\ref{eg:apply GN alg}.

@@

\begin{enumerate}
\item Fix a vertex $i \in [n]$, the {\bf starting vertex}. Set $j:=i$ and $I_i^W = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps:
\begin{enumerate}
\item {\em Step $j$ for starting vertex $i$}: If $\Prop(j) = \emptyset$, do nothing. Else, let $p \in \Prop(j)$ be the clockwise-most propagator supported on $j$. Write $I_i^W = I_i^W\cup \{j\}$, and delete propagator $p$ from the diagram.
\item Increment $j$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

Informally: one starts at $i$ and moves counterclockwise around the vertices of $W$, at each step removing the clockwise-most propagator supported on that vertex (if it exists). The set $I_i^W$ lists the vertices at which a propagator was deleted.

If the algorithm assigns vertex $j$ to propagator $p$ from starting vertex $i$, we say that $p$ \emph{contributes} $j$ to $I_i^W$. Notationally, we represent this by allowing the $I_i^W$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. For each $i \in [n]$, define a function $I_i^W : \cP \longrightarrow [n]$ by
\[I_i^W(p) := \text{the vertex label that $p$ contributes to $I_i$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$. 
\end{dfn}

For both $I_i^W$ the set and the function, when the diagram $W$ is clear from context we drop the superscript and simply write $I_i$.

\begin{dfn} \label{proporderingdfn}
  There is an ordering on the propagators on $W = (\cP, [n])$ defined by $I_i$. This is the ordering in which they contribute to the algorithm as it constructs $I_i$.
%  $\cP = \{p_1, \ldots p_k\}$, where 
% \[I_i = \{I_i(p_1) <_i I_i(p_2) <_i \ldots <_i I_i(p_k)\}.\]
% (The superscript $W$ is omitted here to simplify the notation.)\note{Sian tweaked this slightly to emphasise the $<_i$ order, because standard convention in most papers is to write all GN terms in the 1-order.}
\end{dfn}

Since each propagator is removed by the algorithm after it contributes, each propagator contributed exactly once to $I_i$, making the above ordering well defined.
Note that different $I_*$ impose different orderings on the propagators. Furthermore, Corollary \ref{GN alg well defined} shows that this ordering respects the ordering of vertices in $I_i$. That is, if $p_j$ occurs before $p_l$ in the ordering imposed by $I_i$, then we also have $I_i(p_j) <_i I_i(p_l)$.

@@

To construct $I_1$ we start at vertex 1, so we set $i=1$, $j = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $j = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, $r$ contributes at this step: $I_1(r)=1$. We let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $j = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators, $q$ contributes at this step: $I_1(q)=1$. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $j = 3$: $\Prop(3) = \{p\}$, so now $I_i(p) = 3$ and $I_1 = \{1,2,3\}$. The propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $j = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $j = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$ with $I_1(s) = 5$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. 
%So the propagators are ordered $r,q,p,s$ with respect to $I_1$, and 
$I_1$ viewed as a function is given by
\[I_1: \cP \longrightarrow [n] : \quad I_1(r) = 1,\quad I_1(q) = 2, \quad I_1(p) = 3, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid $M(W)$ associated to the Wilson loop diagram $W$ above. We will prove this in general shortly.
\end{eg}

%
%\begin{dfn} \label{prop order dfn}
%Given a set $\cP$ of $k$ propagators defining a Wilson loop diagram $(\cP, [n])$, the Grassmann necklace element $I_i$ imposes an ordering on it $\cP^i = \{p_1 \ldots p_k\}$ where $p_m$ is the $m^{th}$ propagator to contribute to $I_i$. We write $\cP_m^i = \{p_1 \ldots p_m\}$ to be the first $m$ propagators that contribute to $I_i$. We drop the superscript if the Grassman necklace element is clear.
%\end{dfn}

@@

\begin{figure}
\[
\begin{tikzpicture}[baseline=(current bounding box.east)]
  \begin{scope}
  W' = \drawWLDfragment[12]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \newnode[left]{1}{i}
  \newnode[below]{4}{j}
  \newnode{5}{}
  \newnode{6}{}
  \newnode{7}{}
  \newnode{8}{}
  \newnode{9}{}
  \newnode{11}{}
  \newnode[right]{12}{l+1}
  \begin{scope}
    \clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
    \newpropbend{5.3}{7.6}{80}
    \node at (\zero + 6.5*\step:\radius*0.8) {\footnotesize $r$};
    \newprop[midway,left]{8.5}{-4}{}
    \node at (\zero + 8.15*\step:\radius*0.5) {\footnotesize $p$};
    \newprop[midway]{11.5}{-4}{}
    \node at (\zero + 12.2*\step:\radius*0.7) {\footnotesize $q$};
  \end{scope}
\end{scope}
\end{tikzpicture}
\qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.east)]
  V' = \begin{scope}
  \drawWLDfragment[12]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \newnode[left]{2}{m}
  \newnode[below]{4}{j}
  \newnode{5}{}
  \newnode{6}{}
  \newnode{7}{}
  \newnode{8}{}
  \newnode{9}{}
  \newnode{11}{}
  \newnode[right]{12}{l+1}
  \begin{scope}
    \clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
    \newpropbend{5.3}{7.6}{80}
    \node at (\zero + 6.5*\step:\radius*0.8) {\footnotesize $r$};
    \newprop[midway,left]{8.5}{22}{}
    \node at (\zero + 8.5*\step:\radius*0.5) {\footnotesize $s$};
    \newprop[midway]{11.5}{21}{}
    \node at (\zero + 11.8*\step:\radius*0.8) {\footnotesize $t$};
  \end{scope}
\end{scope}
\end{tikzpicture}
\]
\caption{$W'$ and $V'$ are locally the same between $j$ and $l+1$, so we will have $I_i^W\cap [j,l] = I_m^V\cap[j,l]$ even though $p \neq s$ and $q \neq t$. Note that propagators such as $r$ which have {\em both} ends inside the region of interest must be identical in both diagrams.}\label{fig:locally identical}
\end{figure}

@@

With that in mind, we are ready to begin proving properties of the algorithm. We first show that the algorithm for $I_i$ can never reach the final vertex (in the $<_i$ order) in a propagator's support without having already assigned that propagator; in particular, this guarantees that the construction of $I_i$ always terminates in fewer than $n$ steps.

@@

	\end{scope}
		\end{scope}
	\end{tikzpicture}
\] 
\caption{The two types of configuration that would (in theory) allow the propagator $p$ to contribute its $<_i$-maximal support vertex to $I_i$.}
\label{fig:no fourth vertex}
\end{figure}
In Case 1, \st{observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$.}{\color{violet} since $b+1<_i I_i(p)$ there must be some propagator, $q$ before $p$ in $\cP^i$ such that $I_i(q) = b+1$. By remark \ref{clockwise ordering rem}, $p$ is inside $q$ in the $<_b$ order. Specifically, $q$ must also end on the $b^{th}$ edge. By non-crossing and desnity arguements of admissibility $i-1$ cannot be in $V(q)$.} Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of Case 2. Therefore it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

So let $p = (a,b)$ with $a<_ib$ and $I_i(p) = b+1$, and choose $p$ such that $\big|[a+1,b]\big|$ is minimal amongst propagators of $W$ which contribute their last vertex and are in Case 2.

@@

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators and $n$ vertices, then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ distinct vertices to each $I_i$ in $n$ steps.
\end{cor} 
\begin{proof}
Fix $i \in [n]$. It follows from the proof of Lemma~\ref{lem no fourth vertex} that in applying Algorithm~\ref{alg:put GN on WLD} to construct the set $I_i$, the algorithm can never reach the fourth vertex of a propagator's support (ordered with respect to $<_i$) without having already assigned that propagator. Therefore if the algorithm starts at vertex $i$, all propagators must have been eliminated by the time it gets to vertex $i-1$, ensuring that the algorithm is completed in less than one full circumnavigation of the diagram and that $I_i$ contains exactly $k$ distinct elements.
\end{proof}

@@

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range \hlfix{(since $p$ will be the final propagator encountered by the algorithm)}{Sian added this: it's not really a locally-the-same argument so I'm not sure it'll help to reference Remark 3.12} so if $r \neq p$, then $I_w^W(r) = I_w^V(r)$. Meanwhile, the value of $I^{W}_w(p)$ depends on which step in the algorithm the propagator $q$ contributes a vertex in the diagram $V$, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then 
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority 
%$w \in [l, l+a-1]$} 
%I left this as a comment so we can add it in if necessary, but I don't think we need this level of precision in this sentence: we're just signposting what we're going to do in detail shortly.
of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$ (recall Remark~\ref{rmk algorithm locally same}).  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in Case 2 of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is a Grassmann necklace.
\end{thm}
\begin{proof}
%\note{Sian replaced ``counterclockwise'' with ``clockwise'' in a couple of places in this proof: someone please doublecheck?  Karen checked and it seems correct with Sian's changes.}
% Thank you! Clocks are difficult.
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. We prove that $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is equivalent to the definition of Grassmann necklace given in Definition~\ref{def:grassmann necklace}.

Fix $n$, and suppose for a contradiction that there exists an admissible diagram for which there exists an $i$ with $m\in I_i\setminus \{i\}$ and $m \not\in I_{i+1}$.  Let $W$ be such a counterexample with the minimal number of propagators.

If $i \not\in I_i$, then there are no propagators supported on $i$ at all.  In this case it is clear that applying Algorithm~\ref{alg:put GN on WLD} at vertex $i$ and vertex $i+1$ produces exactly the same result, i.e. $I_{i+1} = I_i$, and so $W$ is not a counterexample at all.

@@

\note{Sian rewrote Case 1 based on Susama's suggested edits}
\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$. By Remark~\ref{rmk algorithm locally same} the construction of $I_i^W$ and $I_{i+1}^W$ is locally the same (indeed, locally {\em identical} in this case) from vertex $i+1$ to $b-1$, so we have $I_{i+1}^W \cap [i+1,b-1] = I_{i}^W \cap [i+1,b-1]$. Note that if $b = i+1$ then this interval is empty and the equality remains true.

%If $p$ is not supported on $i+1$, then in building $I_{i+1}$ we will assign the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, and in particular we have \bas I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]\eas by Remark \ref{rmk algorithm locally same}.  If $p$ is supported on $i+1$, then $b= i+1$ and the interval $[i+1,b-1]$ is empty, so the above equality remains true, albeit vacuously.

By Lemma~\ref{vertex cyclic int lem} it must happen that we have $I_{i+1}^W(p) = b$, as otherwise $b$ would never be contributed by $p$.  Consequently, when the construction of $I_{i+1}^W$ reaches vertex $b$ there cannot be any remaining unassigned propagators that are supported on $b$ and clockwise of $p$ (Remark~\ref{clockwise ordering rem}). It follows that the same must be true when $I_i^W$ reaches vertex $b$, since $I_i^W$ and $I_{i+1}^W$ behave identically on the interval $[i+1,b-1]$ as noted in the previous paragraph.

The previous two paragraphs also imply that $m \geq_i b+1$, since $I_{i+1}^W \cap [i+1,b-1] = I_{i}^W \cap [i+1,b-1]$ and $b \in I_{i+1}^W$.

%This is also true in when the algorithm constructing $I_i$ reaches $b$, \st{since the same propagators have been assigned beforehand.}\sanote{don't like same here}{\color{violet}since any propagators lying counterclockwise of $p$ on the edge $i-1$ must have their support in $[i-1, b]$. By Lemma \ref{lem no fourth vertex}, these propagators contribute before the algorithm reaches $b$. Furthermore, Remark  \ref{rmk algorithm locally same} implies that these propagators contribute the same vertices to both $I_i$ and $I_{i+1}$.} This also implies that $m\geq_i b+1$.

Now let $W' = (\cP_{out}(p,<_i), [n])$, i.e. the diagram obtained from $W$ by removing both $p$ and all propagators inside of $p$. $W'$ has strictly fewer propagators than $W$ did.
